---
title: "Regression Models Course Project"
author: "Wei Lin"
date: "October 2, 2017"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,fig.align = "center",
                      fig.height = 3,fig.width = 5)
library(ggplot2)
library(dplyr)
```

# Executive Summary

1. Manual transmission is better for MPG than automatic ones.

2. Manual tranmission has 2.9358 more MPG than automatic ones.

# Analysis

In mtcars dataset, `am == 0` stands for automatic transmission and `am == 1` stands for manual transmission.

First of we plot a box plot to show the relationship between MPG and transmission, which shows automatic transmission generally has lower mpg than m
```{r}
mtcars_new <- mtcars %>% 
  mutate(transmission = ifelse(am == 0, 'Automatic', 'Manual')) %>% 
  select(-am)
ggplot(mtcars_new) + 
  geom_boxplot(aes(x = transmission, y = mpg)) +
  xlab('Cylinder') + 
  ggtitle('Box Plot of MPG by Cylinder and Transmission')
```

We can fit a linear model of `mpg` by `transmission`. 
```{r}
fit <- lm(mpg ~ transmission, mtcars_new)
coef(summary(fit))
```
In the fitted model, transmission has a coefficient of 7.245, which means cars with automatic tranmission has 7.245 better mpg than manual ones. P-value of the regressor shows that the relationship is significant. The fitted value and residual plot is shown below. There is no clear pattern left in the residual.
```{r}
ggplot() + geom_point(aes(x = fit$fitted.values, y = fit$residuals)) + 
  xlab('Fitted Values') + ylab('Residuals') + 
  geom_hline(aes(yintercept = 0), col = 'red')
```

Nevertheless, if we calculate the correlation matrix, we can find that the variables are highly correlated with each other in the dataset.
```{r}
print(cor(mtcars))
```

Therefore, this simple model does not tell the full story. If we fit a full mode, the coefficient of transmission changes and is not significant anymore. 
```{r}
summary(lm(mpg ~ ., data = mtcars_new)) %>% coef
```

To find a reasonable model, we can use `step` function to find the best model for explaining `mpg`
```{r}
best_fit <- step(lm(mpg ~ ., data = mtcars_new), trace = 0)
summary(best_fit) %>% coef
```
Transmission has a coefficient of 2.9358, which means cars with automatic tranmission has 2.9358 better mpg than manual ones.

The residual plot is shown below. There is no clear pattern left in the residual.
```{r}
ggplot() + geom_point(aes(x = best_fit$fitted.values, 
                          y = best_fit$residuals)) + 
  xlab('Fitted Values') + ylab('Residuals') + 
  geom_hline(aes(yintercept = 0), col = 'red')
```

If we do an anova with and without transmission for the selected model,
we can see the inclusion of transmission is statistically significant.
```{r}
anova(lm(mpg ~ wt + qsec, data = mtcars), best_fit)
```

