# Storm Type Impact Analysis

## Synopsis

We found that the most harmful event type to both health and economics is tornado

## Data Processing
```{r,echo=FALSE,message=FALSE,warning=FALSE}
library(data.table)
library(ggplot2)
library(dplyr)
```

```{r}
knitr::opts_chunk$set(echo = TRUE,message = FALSE,warning = FALSE,
                      fig.align = "center",fig.height = 6,fig.width = 8)
storm_data <- fread('repdata%2Fdata%2FStormData.csv')
```


## Result

### Question 1

Across the United States, which types of events (as indicated in the EVTYPE variable) are most harmful with respect to population health?

Sum fatalities (`FATALITIES`) and injuries (`INJURIES`) to represent the damage to populuation health
```{r}
health_harm_by_type <- storm_data[, .(health_harm = sum(FATALITIES + INJURIES)), EVTYPE]

ggplot(health_harm_by_type %>% head(10)) + geom_bar(aes(x = EVTYPE, weight = health_harm)) + 
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    ggtitle('Top 10 Harmful Events in Health')
```

The most harmful one is `r health_harm_by_type[health_harm == max(health_harm), EVTYPE]`

### Question 2

Across the United States, which types of events have the greatest economic consequences?

Summing property damage (`PROPDMG`) and crop damage (`CROPDMG`) to represent the damage to economics
```{r}
econ_harm_by_type <- storm_data[, .(econ_harm = sum(PROPDMG + CROPDMG)), EVTYPE]
econ_harm_by_type <- econ_harm_by_type[order(econ_harm,decreasing = T)]
ggplot(econ_harm_by_type %>% head(10)) + geom_bar(aes(x = EVTYPE, weight = econ_harm)) + theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    ggtitle('Top 10 Harmful Events in Economics') + 
    ylab('$')
```

The most harmful one is `r econ_harm_by_type[econ_harm == max(econ_harm), EVTYPE]`
