---
title: "Basic Inferential Data Analysis"
author: "Wei Lin"
date: "March 25, 2017"
output:
  pdf_document: default
  html_document: default
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE,warning = FALSE,message = FALSE,
                      fig.height = 5)
```

## Load Data and Summary
```{r}
library(datasets)
dim(ToothGrowth)
head(ToothGrowth)
summary(ToothGrowth)
str(ToothGrowth)
```

## Test

For dose 0.5, we compare the tooth growth by supp using t test where the null hypothesis is that the tooth growth means of the two supp types is equal to zero
```{r}
t.test(len ~ supp, ToothGrowth,subset = ToothGrowth$dose == 0.5)
```
Given low p-value (less than 0.05), we reject the null hypothesis and claim that tooth grow faster under OJ than VC when does is 0.5 


Perform the same test for dose 1
```{r}
t.test(len ~ supp, ToothGrowth,subset = ToothGrowth$dose == 1)
```
Given low p-value (less than 0.05), we reject the null hypothesis and claim that tooth grow faster under OJ than VC when does is 1


Perform the same test for dose 2
```{r}
t.test(len ~ supp, ToothGrowth,subset = ToothGrowth$dose == 2)
```
Given high p-value (higher than 0.05), we cannot reject the null hypothesis and claim that speed of tooth grow is the same under OJ and VC when does is 2

## Conclusion

The OJ makes tooth growth faster than CV when dose is low. However, the benefit diminishes when dose is high.
