---
title: "Central Limit Theorem Simulation for Exponential Distribution"
author: "Wei Lin"
date: "March 25, 2017"
output:
  pdf_document: default
  html_document: default
---

## Overview

We are going to verify the the central limit theorem by simulating a number of exponential distributed random variables.


```{r setup}
knitr::opts_chunk$set(echo = TRUE,warning = FALSE,message = FALSE,
                      fig.height = 3)
set.seed(10)
```

## Simulations

We generation a list of 40 vectors consisting of 1000 exponential random variables
```{r}
random_exp <- sapply(1:1000, function(x) rexp(40, 0.2))
```  
 
## Sample Mean vs Theoretical Mean

The sample mean is shown in red and theoretical mean is shown in black
```{r}       
library(ggplot2)
sample_mean <- apply(random_exp, 2, mean)
ggplot() + geom_histogram(aes(x = sample_mean)) + 
    geom_vline(aes(xintercept = 1 / 0.2),col = 
                   'red')

```

## Sample Variance vs Theoretical Variance

The sample standard deviation is shown in red and theoretical standard deviation is shown in black
```{r}       
sample_sd <- apply(random_exp, 2, sd)
ggplot() + geom_histogram(aes(x = sample_sd)) + 
    geom_vline(aes(xintercept = 1 / 0.2),col = 
                   'red')

```

## Distribution

Distribution of 1000 random exponentials are compared with 1000 of averages of 40 random exponentials
```{r}     
library(tidyr)
compare <- data.frame(
    random_exponential =  rexp(1000, 0.2),
    sample_mean = apply(random_exp, 2, mean))
compare <- compare %>% gather(type, value)


ggplot(compare) + geom_histogram(
    aes(x = value, fill = type),bins = 100)

```

